
In this section we will present some background theory which is
necessary for the constructions and derivations in the following
sections.
This is mainly standard theory of stochastic processes with some
slight extensions.
The reader may skip this if already familiar with such theory
and use this section as a reference when going into the technical
details in the following sections.

\subsubsection{Notation}

We work with a background probability space $(\Omega, \Sigma_\Omega, \Prob)$.
For a measurable space $(\Cal{X}, \Sigma_{\Cal{X}})$ we denote
the set of probability measures on this space $\Cal{P}(\Sigma_\Cal{X})$ or
simply $\Cal{P}(\Cal{X})$ when the $\sigma$-algebra is unambiguous.
When taking cartesian products $\Cal{X} \times \Cal{Y}$ of measurable spaces
$(\Cal{X}, \Sigma_\Cal{X}), (\Cal{Y}, \Sigma_\Cal{Y})$ we always endow such
with the product $\sigma$-algebra $\Sigma_\Cal{X} \otimes \Sigma_\Cal{Y}$,
unless otherwise specified.
A map $f: \Cal{X} \to \Cal{Y}$ is called $\Sigma_{\Cal{X}}$-$\Sigma_{\Cal{Y}}$
measurable provided $f^{-1}(\Sigma_{\Cal{Y}}) \subseteq \Sigma_{\Cal{X}}$
and we denote the set of such functions $\Cal{M}(\Sigma_{\Cal{X}},
\Sigma_{\Cal{Y}})$.
By a random variable $X$ on $(\Cal{X}, \Sigma_{\Cal{X}})$ mean a
$\Sigma_\Omega$-$\Sigma_{\Cal{X}}$ measurable map.

\subsubsection{Kernels}

\begin{defn}[Probability kernel]
  Let $(\Cal{X}, \Sigma_\Cal{X}), (Y, \Sigma_\Cal{Y})$ be measurable spaces.
  A function
  \[ \kappa(\cdot \mid \cdot) : \Sigma_\Cal{Y} \times \Cal{X} \to [0,1] \]
  is a $(\Cal{X}, \Sigma_\Cal{X})$-\defemph{probability kernel}
  on $(\Cal{Y}, \Sigma_\Cal{Y})$ provided
  \begin{enumerate}
    \item $B \mapsto \kappa(B \mid x) \in \Cal{P}(\Sigma_\Cal{Y})$
      that is $\kappa(\cdot \mid x)$ is a probability measure
      for any $x \in \Cal{X}$.
    \item
      $x \mapsto \kappa(B \mid x) \in \Cal{M}(\Sigma_\Cal{X}, \Sigma_\Cal{Y})$
      that is $\kappa(B \mid \cdot)$ is ($\Sigma_\Cal{X}$-$\Sigma_\Cal{Y}$)
      measurable for any $B \in \Sigma_\Cal{Y}$.
  \end{enumerate}
  When the $\sigma$-algebras are unambiguous we shall simply say an
  $\Cal{X} \leadsto \Cal{Y}$ kernel.
  For any $x \in \Cal{X}$ and $f \in \Cal{L}_1(\kappa(\cdot \mid x))$
  we write the integral of $f$ over $\kappa(\cdot \mid x)$ as
  $\int f(y) \difd \kappa(y \mid x)$.
  \label{defn:probKer}
\end{defn}

We now state some fundamental results on probability kernels
\begin{thm}[Integration of a kernel]
  Let $\mu \in \Cal{P}(\Cal{X})$ and $\kappa : \Cal{X} \leadsto \Cal{Y}$.
  Then there exists a uniquely determined probability measure
  $\lambda \in \Cal{P}(\Sigma_\Cal{X} \otimes \Sigma_\Cal{Y})$
  such that
  \[ \lambda(A \times B) = \int_A \kappa(B, x) \difd \mu(x) \]
  \label{thm:intKer}
  We denote this measure $\lambda = \kappa \mu$.
\end{thm}
\begin{proof}
  We refer to \ncite{RH14} thm. 1.2.1.
\end{proof}

Notice that by \cref{thm:intKer}
besides getting a probability measure on $\Cal{X} \times \Cal{Y}$
we get an induced probability measure
on $\Cal{Y}$ defined by $B \mapsto (\kappa \mu)(\Cal{X} \times B)$.
We will denote this measure by $\kappa \circ \mu$.
This way $\kappa$ can also be seen as a mapping from
$\Cal{P}(\Cal{X}) \to \Cal{P}(\Cal{Y})$.
Also note that $\kappa \circ \delta_x = \kappa(\cdot \mid x)$.

For an idea how to actually compute integrals over kernel derived measures
we here include
\begin{thm}[Extended Tonelli and Fubini]
  Let $\mu \in \Cal{P}(\Cal{X})$,
  $f \in \Cal{M}(\Sigma_\Cal{X} \otimes \Sigma_\Cal{Y}, \bb{B})$
  be a measurable function and
  $\kappa : \Cal{X} \leadsto \Cal{Y}$ be a probability kernel.
  Then
  \[ \int \abs{f} \difd \kappa \mu
  = \int \int \abs{f} \difd \kappa(\cdot \mid x) \difd \mu(x) \]
  Furthermore if this is finite, i.e. $f \in \Cal{L}_1(\kappa(\cdot, \mu))$
  then $A_0 \defeq \left\{ x \in \Cal{X} \Mid
    \int f \difd \kappa(\cdot \mid x) < \infty \right\}
  \in \Sigma_\Cal{X}$
  with $\mu(A_0) = 1$, 
  \[ x \mapsto \begin{cases}
      \int f \difd \kappa(\cdot \mid x) & x \in A_0
      \\ 0 & x \not\in A_0
  \end{cases} \]
  is $\Sigma_\Cal{X}$-$\bb{B}$ measurable and
  \[ \int f \difd \kappa \mu
  = \int_{A_0} \int f \difd \kappa(\cdot \mid x) \difd \mu(x) \]  
  \label{thm:extTonFub}
\end{thm}
\begin{proof}
  We refer to \ncite{RH14} thm. 1.3.2 and 1.3.3.
\end{proof}

\begin{prop}[Composition of kernels]
  Let $\kappa : \Cal{X} \leadsto \Cal{Y}, \psi : \Cal{Y} \leadsto \Cal{Z}$
  be probability kernels. Then
  \[ (\psi \circ \kappa)(A \mid x) \defeq
    \int \psi(A \mid y) \difd \kappa(y \mid x)
  ,\qquad \forall A \in \Sigma_{\Cal{Z}}, x \in \Cal{X} \]
  is a $\Cal{X} \leadsto \Cal{Z}$ probability kernel called the
  composition of $\kappa$ and $\psi$. The composition operator
  $\circ$ is associative, i.e. if $\phi : \Cal{Z} \leadsto \Cal{W}$ is
  a third probability kernel then $(\phi \circ \psi) \circ \kappa = 
  \phi \circ (\psi \circ \kappa)$.
  The associativity also extends to measures, i.e.
  $\forall \mu \in \Cal{X}
  : (\psi \circ \kappa) \circ \mu = \psi \circ (\kappa \circ \mu) $
  and this is uniquely determined by $\psi, \kappa$ and $\mu$.
  \label{prop:compKer}
\end{prop}
\begin{proof}
  The first assertion is a trivial verification of the two conditions
  in \cref{defn:probKer} and left as an exercise.
  For the associativity we refer to \ncite{RH14} lem. 4.5.4.
\end{proof}

\Cref{prop:compKer} actually makes the class of measurable spaces
into a category \mcite{L62},
with identity $\id_{\Cal{X}}(\cdot \mid x) = \delta_x$.
Notice that the mapping $(A\times B, x) \mapsto \delta_x(A) \kappa(B \mid x)$
defines a probability kernel $\Cal{X} \leadsto \Cal{X} \times \Cal{Y}$
which we could denote $\id_{\Cal{X}} \times \kappa$.
Now if $\psi : \Cal{X} \times \Cal{Y} \leadsto \Cal{Z}$ is a kernel
then by \cref{prop:compKer} the composition
$(\id_{\Cal{X} \times \Cal{Y}} \times \psi)
\circ (\id_{\Cal{X}} \times \kappa)$
is a kernel $\Cal{X} \to \Cal{X} \times \Cal{Y} \times \Cal{Z}$
which we will denote $\psi \kappa$.
It inherits associativity from $\circ$ and again this associativity
extends to application on measures: if $\mu$ is a measure on $\Cal{X}$
then $\psi (\kappa \mu) = (\psi \kappa) \mu$.

\begin{prop}
  Let $\kappa : \Cal{X} \to \Cal{Y}$ be a probability kernel
  and $f : \Cal{Y} \to \ol{\ul{\R}}$ be integrabel.
  Then $x \mapsto \int f \difd \kappa(\cdot \mid x)$ is measurable
  into $(\ol{\ul{\R}}, \ol{\ul{\bb{B}}})$.
  \label{prop:intKerMeas}
\end{prop}
\begin{proof}
  Simple functions are measurable since $\kappa$ is a kernel.
  Now extend by sums and limits.
\end{proof}

\subsubsection{Kernel derived processes}

Let $(\Cal{X}_n, \Sigma_{\Cal{X}_n})_{n \in \N}$ be a sequence
of measurable spaces. For each $n \in \N$ define
$\Cal{X}^{\ul{n}} \defeq \Cal{X}_1 \times \dots \times \Cal{X}_n$,
$\Sigma_{\Cal{X}^{\ul{n}}} \defeq \Sigma_{\Cal{X}_1} \otimes
\dots \otimes \Sigma_{\Cal{X}_n}$
and let
$\kappa_n : \Cal{X}^{\ul{n}} \leadsto \Cal{X}_{n+1}$ be a probability kernel.
Then $\kappa^{\ul{n}} \defeq \kappa_n \dots \kappa_1$ is a kernel
from $\Cal{X}_1$ to $\Cal{X}^{\ul{n+1}}$.
So for any probability measure $\rho_1 \in \Cal{P}(\Cal{X}_1)$
there exists a unique probability measure 
$\rho_n$ on $\Cal{X}^{\ul{n}}$ defined by
$\kappa^{\ul{n-1}} \rho_1$.
This however does not imply existence of a measure on
\[ \left(\Cal{X}^{\ul{\infty}},
  \Sigma_{\Cal{X}^{\ul{\infty}}} \right) \]
where $\Cal{X}^{\ul{\infty}} \defeq \prod_{n \in \N} \Cal{X}_n$ and
$\Sigma_{\Cal{X}^{\ul{\infty}}} \defeq \bigotimes_{n \in \N}
\Sigma_{\Cal{X}_n} $.
We will need this to talk about countable stochastic processes
arising from decision models, so we include here a result
by C. Ionescu-Tulcea (1949):

\begin{thm}[Ionescu-Tulcea extension theorem]
  For every $\mu \in \Cal{P}(\Cal{X}_1)$ 
  there exists a unique probability measure
  $\rho \in \Cal{P}(\Cal{X}^{\ul{\infty}})$ such that
  \[ \kappa^{\ul{n-1}} \mu (A) = \rho
    \left( A \times \prod_{k=n+1}^\infty \Cal{X}_k \right)
  , \qquad \forall A \in \Sigma_{\Cal{X}^{\ul{n}}}, n \in \N \]
  \label{thm:ionescuTulcea}
\end{thm}
\begin{proof}
  We refer to \mcite{K02} thm. 5.17. %todo do the proof instead
\end{proof}

We will also need to integrate with respect to starting conditions
on such a stochastic process. To this end it is convenient to
extend the Ionescu-Tulcea construction from a measure to a kernel.
As I was not able to find a proof of this elsewhere
an original proof is included here.
\begin{prop}[Ionescu-Tulcea kernel]
  Let $\mu_x$ denote the Ionescu-Tulcea measure of a
  sequence of probability kernels
  $\kappa_i : \Cal{X}^{\ul{i}} \to \Cal{X}_{i+1}$
  with starting measure $\delta_x$ on $\Cal{X}_1$ for any $x \in \Cal{X}_1$.
  Then $\kappa(A \mid x) = \mu_x(A)$ defines a probability kernel
  $\kappa : \Cal{X}_1 \to \Cal{X}^{\ul{\infty}}$.
\end{prop}
\begin{proof}
  Since we already know that $\mu_x$ is a probability measure for any
  $x \in \Cal{X}_1$,
  we just have to show that $\kappa(A \mid x) = \mu_x(A)$ is measurable
  as a function of $x$ for all
  $A \in \Sigma_{\Cal{X}^{\ul{\infty}}}
  = \bigotimes_{i=1}^\infty \Sigma_{\Cal{X}_i}$.
  Let $\phi_A = x \mapsto \mu_x(A)$
  for all $A \in \Sigma_\Cal{X^{\ul{\infty}}}$ and define
  \[ \bb{G} = \left\{ A \in \bigotimes_{i=1}^\infty \Sigma_{\Cal{X}_i}
  \Mid \phi_A \in \Cal{M}(\Sigma_{\Cal{X}_1}, \bb{B}_{[0,1]}) \right\} \]
  The cylinder algebra
  \[ \bb{C} = \left\{ A_1 \times \dots \times A_i \times \Cal{X}_{i+1},
  \dots \Mid A_i \in \Sigma_{\Cal{X}_i}, i \in \N \right\} \]
  is a generator for $\Sigma_{\Cal{X}^{\ul{\infty}}}$ stable under 
  finite intersections.
  By contruction $\bb{C} \subseteq \bb{G}$ since
  \[ \phi_{A_1 \times \dots \times A_i \times \Cal{X}_{i+1} \times \dots}
  = \kappa^{\ul{i-1}}(A_1 \times \dots \times A_i \mid \cdot) \]
  and any $\kappa^{\ul{i-1}}$ is a kernel (\cref{prop:compKer}).
  We will show that $\bb{G}$ is a Dynkin class.
  Then by Dynkins $\pi$-$\lambda$ theorem (see \cref{thm:DynkinPiLambda})
  \[ \sigma(\bb{C}) = \Sigma_{\Cal{X}^{\ul{\infty}}}
  \subseteq \bb{G} \]
  implying that $\phi_A$ is measurable
  for all $A \in \Sigma_{\Cal{X}^{\ul{\infty}}}$.
  
  Clearly $\Cal{X}^{\ul{\infty}}, \emptyset \in \bb{G}$ and if
  $A,B \in \bb{G}$ with $A \subseteq B$ then
  $\phi_{B \setminus A} = \phi_B - \phi_A \in \bb{G}$.
  Finally if $(B_n)_{n \in \N}$ is an ($\subseteq$-) increasing sequence
  in $\bb{G}$ then $\phi_{\bigcup_{n=1}^\infty B_n} =
  \lim_{n \to \infty} \phi_{B_n}$ is again measurable as it is a
  limit of measurable functions, showing that $\bb{G}$ is a Dynkin class.
\end{proof}

We will denote the Ionescu-Tulcea kernel $\dots \kappa_2 \kappa_1$ or
$\prod_{i=1}^\infty \kappa_i$ or simply $\kappa^{\ul{\infty}}$.

\begin{lem}
  The Ionescu-Tulcea kernel satisfies
  $\prod_{i=1}^\infty \kappa_i = (\prod_{i=2}^\infty \kappa_i) \kappa_1 $.
  \label{lem:ionescu}
\end{lem}
\begin{proof}
  Let $x \in \Cal{X}_1$.
  Notice that by associativity of the finitely induced measures
  $\kappa_n \dots \kappa_1 \delta_x
  = (\kappa_n \dots \kappa_2) (\kappa_1 \delta_x)$.
  This implies that
  \[ \prod_{i=1}^\infty \kappa_i \delta_x
    \left( A \times \prod_{k=n+1}^\infty \Cal{X}_k \right)
    = \prod_{i=2}^\infty \kappa_i \kappa_1 \delta_x
  \left( A \times \prod_{k=n+1}^\infty \Cal{X}_k \right) \]
  for all $n \in \N$ and $A \in \Sigma_{\Cal{X}^{\ul{n}}}$.
  By the uniqueness in \cref{thm:ionescuTulcea} we are done.
\end{proof}
This lemma will come in handy when manipulating with integrals over
kernel derived measures.

